---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

Welcome to my Titanic notebook. I have two Points to try here:

1. Use the floor-plans to get actual 3D-Information about the Cabins
2. Use increasingly more accurate Bayesian Modelling, to see if it helps.

But first I'll build a baseline, that incorporate all the "usual" features, used by successful participants. Most of them I found myself, but the Name-Processing was something I completely missed.
Full credit for that goes to all the other participants.

```{r}

library(brms) # Bayesian
library(mgcv) # Non-Bayesian
library(parallelly)
library(ggplot2)
library(dplyr)

options(
  mc.cores = availableCores()
)

#library(future)
#plan(multisession, workers = 5)

```

# Model Evaluation

We'll evaluate all Modes with 5 fold cross validation.

```{r}
evaluate_brms_model <- function(fit) {
  # fit_loo <- loo(fit, moment_match = TRUE)
  fit_kfold <- kfold(fit, chains = 1, save_fits = TRUE, folds = all_df[ITrain,"kfold"])
  fit_preds <- kfold_predict(fit_kfold)
  fit_accuracy <- mean(fit_preds$y == ifelse(colMeans(fit_preds$yrep) >= 0.5, 1, 0))
  fit_full_accuracy <- mean(fit$data$Survived == ifelse(predict(fit)[,"Estimate"] >= 0.5, 1, 0))
  
  rbind(
    #fit_loo$estimates, 
    fit_kfold$estimates, data.frame(Estimate=fit_accuracy, SE=NA, row.names = "accuracy_kfold"), data.frame(Estimate=fit_full_accuracy, SE=NA, row.names = "accuracy"))
}


evaluate_gam_formula <- function(formula, data) {
  train_df <- data[ITrain, ]
  K <- max(train_df$kfold)
  cv_pred <- vector("list", K)
  cv_score <- numeric(K)
  cv_score_mse <- numeric(K)
  for (k in seq_len(K)) {
    train_k <- train_df[train_df$kfold != k, ]
    test_k  <- train_df[train_df$kfold == k, ]
  
    fit <- gam(
      formula,
      data = train_k,
      family = binomial(link = "logit")
    )
  
    pred <- predict(fit, newdata = test_k, type = "response")
    
    cv_pred[[k]] <- ifelse(pred >= 0.5, 1, 0)
    cv_score[k] <- mean(test_k$Survived == cv_pred[[k]])  # Accuracy
    cv_score_mse[k] <- mean((test_k$Survived - pred)^2)  # MSE
  }
  
  fit <- gam(
    formula,
    data = train_df,
    family = binomial(link = "logit")
  )
  fit_full_accuracy <- mean(fit$model$Survived == ifelse(inv_logit_scaled(predict(fit)) >= 0.5, 1, 0))
  
  formula_str <- deparse(formula, width.cutoff=500, nlines = 1)
  rbind(
    data.frame(formula=formula_str, Acc=fit_full_accuracy, KMeansAcc=mean(cv_score), KMeansAccSe=sd(cv_score), KMeansMse=mean(cv_score_mse), KMeansMseSe=sd(cv_score_mse))
  )
}

write_gam_solution <- function(formula, full_data, filename) {
  fit <- gam(
    formula,
    data = full_data[ITrain,],
    family = binomial(link = "logit")
  )
  fit_prediction <- predict(fit, newdata=full_data[ITest,])
  fit_solution <- data.frame(PassengerId = full_data[ITest,"PassengerId"], Survived=(fit_prediction >= 0.5) * 1)
  
  write.csv(fit_solution, file = filename, row.names = FALSE)
}


predict_mean <- function(fit) {
  if (inherits(fit, "brmsfit")) {
    # brms: posterior mean on response scale
    preds <- fitted(fit, scale = "response")
    return(preds[, "Estimate"])
  }
  if (inherits(fit, "gam")) {
    # mgcv: response scale
    return(predict(fit, type = "response"))
  }
  # --- projpred projection ---
  if (inherits(fit, "projection")) {
    return( colMeans(proj_linpred(fit, transform=TRUE)$pred) )
  }
  stop("Unsupported model class")
}
model_fit_data <- function(fit) {
  if (inherits(fit, "brmsfit")) {
    return(fit_refit$data)
  }
  if (inherits(fit, "gam")) {
    return(fit$model)
  }
  if (inherits(fit, "projection")) {
    return( fit$refmodel$fetch_data() )
  }
  stop("Unsupported model class")
}

analyze_fit <- function(fit) {
  data_df <- cbind(model_fit_data(fit), Prediction=predict_mean(fit)) %>%
    mutate(
      Survived = factor(Survived > 0.5, levels = c(TRUE, FALSE)),
      PredictionBin = factor(Prediction > 0.5, levels = c(FALSE, TRUE))
    ) 
  
  plot_df <- data_df %>% count(Survived, PredictionBin)
  
  accuracy <- mean(data_df$Survived == data_df$PredictionBin)
  likelihood <- 0 #logLik(fit) # funzt nicht für projected
  
  library(pROC)
  roc_obj <- roc(response = data_df$Survived, predictor = data_df$Prediction)
  
  print(
    plot(
      roc_obj,
      col = "#1565c0",
      lwd = 3,
      main = sprintf("ROC curve (AUC = %.3f)", auc(roc_obj)),
      sub = deparse(fit$formula)
    )
    #abline(a = 0, b = 1, lty = 2, col = "grey")
  )
  
  ggplot(plot_df, aes(x=Survived, y=PredictionBin, fill=n)) + geom_tile() + geom_text(aes(label = n), size = 6, fontface = "bold") +
    scale_fill_gradient(low = "#e3f2fd", high = "#1565c0") +   theme_minimal(base_size = 13) + 
    labs( x = "True outcome (Survived)", y = "Predicted outcome",
          title = paste0("Confusion Matrix, ", round(100*accuracy,1), "% Acc, ", round(likelihood), " l.Lik. ", round(auc(roc_obj)*100,1)," AUC"), subtitle = deparse(fit$formula))
}

scale_df <- function(df) {
  df %>%
    mutate(
      across(
        where(is.numeric) & !all_of("Survived"),
        ~ as.numeric(scale(.x))
      )
    )
}
```

# Feature Engineering

I'm putting Test- and Training Data in one DF, so I can do all the feature Engineering in a simple way.
Unknown Values are assigned "NA". All categorical data is turned into a factor.

```{r}
all_df <- rbind(read.csv2("data/train.csv", sep = ","), 
                cbind(read.csv2("data/test.csv", sep = ","), Survived=NA))

NTrain <- sum(!is.na(all_df$Survived))
ITrain <- seq(1,NTrain)
ITest <- seq(NTrain+1,nrow(all_df))

all_df$kfold <- rep(1:5, 300)[1:nrow(all_df)]

all_df$Age <- as.numeric(all_df$Age)
all_df$Fare <- as.numeric(all_df$Fare)
all_df$SibSp <- as.numeric(all_df$SibSp)
all_df$Parch <- as.numeric(all_df$Parch)

embarked_levels <- c("C", "Q", "S")
all_df$Embarked <- factor(all_df$Embarked, levels=embarked_levels)

sex_levels <- c("male", "female")
all_df$Sex <- factor(all_df$Sex, levels=sex_levels)

pclass_levels <- c("First", "Second", "Third")
all_df$Pclass <- factor(all_df$Pclass, labels = pclass_levels)

rbind(
  evaluate_gam_formula(Survived ~ 1, all_df),
  evaluate_gam_formula(Survived ~ Sex, all_df),
  evaluate_gam_formula(Survived ~ Pclass, all_df)
)
```

## Names

As others have shown, the Title in the Name is very important for a good fit.
Some also use the Surname as important feature. Therefore I'm including that as well.

```{r}
# All Titles, that exist in the training data at least two times
title_levels <- c("Other", "Mr", "Miss", "Mrs", "Master", "Dr", "Rev", "Col", "Major", "Mlle")
all_df$Title <- substring(all_df$Name,regexpr(",",all_df$Name)+2,regexpr("\\.",all_df$Name)-1)
all_df$Title <- factor(all_df$Title, levels=title_levels)
all_df$Title[is.na(all_df$Title)] <- "Other"

# Careful to not use Test-Data, here as well
all_df$Surname <- substring(all_df[, c("Name")],0,regexpr(",",all_df[, c("Name")])-1)
all_df$Surname <- gsub("[ '\\-]", "_", all_df$Surname)

surname_levels <- all_df$Surname[ITrain]
surname_levels <- unique(surname_levels[ ave(ITrain, surname_levels, FUN=length) > 1 ])
surname_levels <- c("Other", surname_levels)

all_df$Surname <- factor(all_df$Surname, levels=surname_levels)
all_df$Surname[is.na(all_df$Surname)] <- "Other"

rbind(
  evaluate_gam_formula(Survived ~ Title, all_df),
  evaluate_gam_formula(Survived ~ Title*Sex, all_df)
)
# Surname schwierig, weil bei der cross-validation eingie Werte wegfallen
# evaluate_gam_formula(Survived ~ Surname, all_df)
```

## Ticket String

```{r}
ticketPrefix_levels <- substring(all_df$Ticket,1,regexpr("[ ]",all_df$Ticket)-1)
ticketPrefix_levels <- gsub("[/.]", "", ticketPrefix_levels)

all_df$TicketPrefix <- ticketPrefix_levels

ticketPrefix_levels <- ticketPrefix_levels[ITrain]
ticketPrefix_levels <- unique(c("", "Other", ticketPrefix_levels[ ave(ITrain, ticketPrefix_levels, FUN=length) >= 1 ]))

all_df$TicketPrefix <- factor(all_df$TicketPrefix, levels = ticketPrefix_levels)
all_df$TicketPrefix[is.na(all_df$TicketPrefix)] <- "Other"

# TicketPrefix schwierig, weil bei der cross-validation eingie Werte wegfallen
# evaluate_gam_formula(Survived ~ TicketPrefix, all_df)
```

## Cabin String

Sometimes we have multiple Cabins in one Strings. In these cases, the first one sometimes is missing a Number. We get rid by all that complexity by always taking the last one. The evenness can be interpreted as the ship's side.

```{r}
all_df$LastCabin <- substring(all_df$Cabin,regexpr(" [^ ]+$",all_df$Cabin)+1,length(all_df$Cabin)-1)
all_df$CabinUnknown <- as.numeric(all_df$Cabin == "")

deck_levels <- c("A", "B", "C", "D", "E", "F", "G", "T")
all_df$Deck <- factor(substr(all_df$LastCabin,0,1), levels=deck_levels)
all_df$CabinNr <- as.numeric(substring(all_df$LastCabin,2))
all_df$CabinSide <- (all_df$CabinNr %% 2) * 2

all_df$DeckOrUnknown <- factor(all_df$Deck, levels=c("Unknown", deck_levels))
all_df$DeckOrUnknown[is.na(all_df$DeckOrUnknown)] <- "Unknown"
all_df$CabinSideOrZero <- all_df$CabinSide
all_df$CabinSideOrZero[is.na(all_df$CabinSide)] <- 1

# Deck schwierig, weil bei der cross-validation eingie Werte wegfallen
# evaluate_gam_formula(Survived ~ Deck, all_df)
rbind(
  evaluate_gam_formula(Survived ~ CabinNr + CabinSide, all_df),
  evaluate_gam_formula(Survived ~ factor(CabinSideOrZero), all_df)
)
  
```

## Embarked Imputation

```{r}
all_df$Embarked[is.na(all_df$Embarked)] <- "S"

evaluate_gam_formula(Survived ~ Embarked, all_df)
```

## Fare Imputation

```{r}
fare_fit <- lm(formula = log(Fare+0.0001) ~ 0 + Pclass + SibSp + Parch + Embarked, data = all_df[ITrain,])

all_df$Fare[is.na(all_df$Fare)] <- exp(predict(fare_fit, all_df[is.na(all_df$Fare),]))-0.0001
all_df$LogFare <- log(1+all_df$Fare)

fare_levels <- c("Free", "Cheap", "Regular", "Pricy", "Luxury")
# Values based on Quantiles of training set
all_df$FareBin <- ifelse(all_df$Fare == 0, "Free", ifelse(all_df$Fare <= 7.9, "Cheap", ifelse(all_df$Fare <= 14.5, "Regular", ifelse(all_df$Fare <= 31.3, "Pricy", "Luxury"))))
all_df$FareBin <- factor(all_df$FareBin, levels = fare_levels)

rbind(
  evaluate_gam_formula(Survived ~ Fare, all_df),
  evaluate_gam_formula(Survived ~ LogFare, all_df),
  evaluate_gam_formula(Survived ~ FareBin, all_df)
  )
```

## Age Imputation based on Title

Binning makes sense, numerical age is useless then.

```{r}
age_fit <- lm(formula = log(Age) ~ 0 + Title + Pclass + SibSp + Parch + Fare + Embarked + Sex, data = all_df[ITrain,])

all_df$AgeUnknown <- is.na(all_df$Age)

all_df$AgeImputed <- all_df$Age
all_df$AgeImputed[is.na(all_df$Age)] <- exp(predict(age_fit, all_df[is.na(all_df$Age),]))

age_levels <- c("Infant", "Teenager", "Adult", "Old")
all_df$AgeBin <- ifelse(all_df$AgeImputed <= 6, "Infant", ifelse(all_df$AgeImputed <= 18, "Teenager", ifelse(all_df$AgeImputed <= 60, "Adult", "Old")))
all_df$AgeBin <- factor(all_df$AgeBin, levels = age_levels)

rbind(
  evaluate_gam_formula(Survived ~ AgeImputed, all_df),
  evaluate_gam_formula(Survived ~ AgeBin, all_df),
  evaluate_gam_formula(Survived ~ AgeImputed + AgeBin, all_df)
  )
```


## Family Structure

```{r}
all_df$SoloTraveler <- ifelse(all_df$SibSp + all_df$Parch == 0, 1, 0)
all_df$AdultWithInfant <- as.integer(ave(all_df$AgeBin == "Infant", all_df$Ticket, FUN = any) & all_df$AgeBin == "Adult")
all_df$AdultWithBoy <- as.integer(ave(all_df$AgeBin == "Infant" & all_df$Sex == "male", all_df$Ticket, FUN = any) & all_df$AgeBin == "Adult")
all_df$AdultWithTeenager <- as.integer(ave(all_df$AgeBin == "Teenager", all_df$Ticket, FUN = any) & all_df$AgeBin == "Adult")
all_df$AdultWithOld <- as.integer(ave(all_df$AgeBin == "Old", all_df$Ticket, FUN = any) & all_df$AgeBin == "Adult")
all_df$TicketTravelers <- as.integer(ave(rep(1,nrow(all_df)), all_df$Ticket, FUN = sum)) - 1

all_df$Namesakes <- pmax( as.integer(ave(all_df$Surname != "Other", all_df$Surname, FUN = sum)) - 1, 0)

rbind(
  evaluate_gam_formula(Survived ~ SoloTraveler, all_df),
  evaluate_gam_formula(Survived ~ AdultWithInfant, all_df),
  evaluate_gam_formula(Survived ~ AdultWithTeenager, all_df),
  evaluate_gam_formula(Survived ~ AdultWithOld, all_df),
  evaluate_gam_formula(Survived ~ TicketTravelers, all_df),
  evaluate_gam_formula(Survived ~ Namesakes, all_df),
  evaluate_gam_formula(Survived ~ SoloTraveler + AdultWithInfant + AdultWithTeenager + AdultWithOld + TicketTravelers + Namesakes, all_df)
  )

rbind(
  evaluate_gam_formula(Survived ~ Sex*SoloTraveler, all_df),
  evaluate_gam_formula(Survived ~ Sex*AdultWithInfant, all_df),
  evaluate_gam_formula(Survived ~ Sex*AdultWithTeenager, all_df),
  evaluate_gam_formula(Survived ~ Sex*AdultWithOld, all_df),
  evaluate_gam_formula(Survived ~ Sex*TicketTravelers, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes, all_df),
  evaluate_gam_formula(Survived ~ Sex + SoloTraveler + AdultWithInfant + AdultWithTeenager + AdultWithOld + TicketTravelers + Namesakes, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes, all_df)
  )

rbind(
  evaluate_gam_formula(Survived ~ Sex*Namesakes + SoloTraveler, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes + AdultWithInfant, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes + AdultWithTeenager, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes + AdultWithOld, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes + TicketTravelers, all_df),
  evaluate_gam_formula(Survived ~ Sex*Namesakes + AdultWithInfant + AdultWithTeenager, all_df)
  )
```

# Linear Model Performance Analysis

Let's see where linear models thrive and where they fail with these features.
```{r}

analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ 1))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ AgeImputed))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ s(AgeImputed)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ AgeBin))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ s(AgeImputed, by=Sex)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ AgeBin*Sex))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ AgeBin + Sex))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * AgeUnknown))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * Pclass))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * factor(SibSp)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * factor(Parch)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ s(Fare, by = Sex)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * LogFare))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * FareBin))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * Embarked))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * Title))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * Surname))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * TicketPrefix))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * CabinUnknown))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * DeckOrUnknown))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * factor(CabinSideOrZero)))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * SoloTraveler))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * AdultWithInfant)) # hmm
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * AdultWithBoy))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * AdultWithTeenager))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * AdultWithOld))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * TicketTravelers))
analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), Survived ~ Sex * Namesakes))


formula_full <- Survived ~ Sex+s(AgeImputed, by=Sex)+AgeUnknown+Pclass+factor(SibSp)+factor(Parch)+s(Fare, by = Sex)+LogFare+FareBin+
                      Embarked+Title+Surname+TicketPrefix+CabinUnknown+DeckOrUnknown+factor(CabinSideOrZero)+SoloTraveler+AdultWithInfant+AdultWithBoy+AdultWithTeenager+AdultWithOld+
                      TicketTravelers+Namesakes
formula_full2 <- Survived ~ Sex+s(AgeImputed)+AgeUnknown+Pclass+factor(SibSp)+factor(Parch)+s(Fare)+LogFare+FareBin+
                      Embarked+Title+Surname+TicketPrefix+CabinUnknown+DeckOrUnknown+factor(CabinSideOrZero)+SoloTraveler+AdultWithInfant+AdultWithBoy+AdultWithTeenager+AdultWithOld+
                      TicketTravelers+Namesakes


analyze_fit(gam(data = all_df[ITrain, ], family = binomial(link = "logit"), formula_full))

fit_full <- brm(
  formula_full,
  data = all_df[ITrain, ],
  family = bernoulli(),
  iter = 2000,
  silent = FALSE,
  refresh = 100,
  prior = prior(student_t(3, 0, 5), class = "b"),
  control=list(max_treedepth=20)
)
fit_reg <- brm(
  formula_full,
  data = all_df[ITrain, ],
  family = bernoulli(),
  iter = 2000,
  silent = FALSE,
  refresh = 100,
  prior = prior(horseshoe(), class = "b")
)
fit_reg2 <- brm(
  formula_full2,
  data = all_df[ITrain, ],
  family = bernoulli(),
  iter = 2000,
  silent = FALSE,
  refresh = 100,
  prior = prior(horseshoe(), class = "b")
)

analyze_fit(fit_full)
analyze_fit(fit_reg)

library(projpred)
vs <- varsel( # DAS DAUERT!
  fit_reg2,
  method = "forward",
  nterms_max = 10
)
plot(vs)
size <- suggest_size(vs, stat = "elpd", se = 1)
size
fit_reduced <- project(vs, size = size)
selected_terms <- solution_terms(vs, size = size)
selected_terms
selected_formula <- reformulate(selected_terms, response = "Survived")
fit_refit <- brm(
  selected_terms,
  data = all_df[ITrain,],
  family = bernoulli(),
  prior = prior(horseshoe(), class = "b")
)
fit_refit2 <- brm(
  selected_terms,
  data = all_df[ITrain,],
  family = bernoulli(),
  prior = prior(student_t(3, 0, 5), class = "b") # less regularization
)
fit_refit_gam <- gam(selected_formula, data = scale_df(all_df[ITrain, ]), family = binomial(link = "logit"))
fit_refit_gam2 <- gam(selected_formula, data = scale_df(all_df[ITrain, ]), family = binomial(link = "logit"), select = TRUE)
fit_refit_gam3 <- gam(selected_formula, data = scale_df(all_df[ITrain, ]), family = binomial(link = "logit"), method = "REML")

analyze_fit(fit_reduced)
analyze_fit(fit_refit)
analyze_fit(fit_refit_gam)
analyze_fit(fit_refit_gam2)
analyze_fit(fit_refit_gam3)

```
Compare brms and gam coefficients

```{r}
library(tibble)
brms_coef <- fixef(fit_refit) |>
  as.data.frame() |>
  rownames_to_column("term") |>
  transmute(
    term,
    brms_est = Estimate,
    brms_se  = Est.Error
  )

# --- gam ---
gam_sum <- summary(fit_refit_gam3)

gam_coef <- gam_sum$p.table |>
  as.data.frame() |>
  rownames_to_column("term") |>
  transmute(
    term,
    gam_est = Estimate,
    gam_se  = `Std. Error`
  )

coef_compare <- full_join(brms_coef, gam_coef, by = "term") |>
  mutate(
    diff = brms_est - gam_est,
    abs_diff = abs(diff)
  )

coef_compare

ggplot(coef_compare, aes(x = gam_est, y = brms_est)) +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_text(aes(label = term), hjust = -0.1, size = 3) +
  labs(
    x = "GAM coefficient",
    y = "brms coefficient",
    title = "Coefficient comparison (brms vs GAM)"
  ) +
  theme_minimal()

conditional_effects(fit_refit, "AgeImputed")
plot(fit_refit_gam3, select = 1)
```


# Feature Selection

LASSO-Implementation acc. to https://rpubs.com/Momo2019/1184100

```{r}
# install.packages("glmnet")
library(glmnet)
# install.packages("psych")
library(psych)

# glmnet(as.matrix(all_df[ITrain,c("AgeImputed", "Fare")]), as.matrix(all_df[ITrain,c("Survived")]))

lasso.features = c("Pclass", "AgeImputed", "Fare", "LogFare", "Sex",
                   "SibSp", "Parch", "SoloTraveler", "AdultWithInfant", "AdultWithBoy", "AdultWithTeenager", "AdultWithOld", "TicketTravelers", "Namesakes")
lasso.X <- transform(all_df[,lasso.features], Sex=as.numeric(Sex), Pclass=as.numeric(Pclass))
lasso.y <- all_df[,c("Survived")]

myDummyCode <- function(column) {
  dummyEncoded <- dummy.code(all_df[,column], )
  colnames(dummyEncoded) <- paste0(column, "_", colnames(dummyEncoded))
  dummyEncoded
}

# One-Hot encoded factors
lasso.X <- cbind(
  lasso.X,
  myDummyCode("AgeBin"),
  myDummyCode("FareBin"),
  myDummyCode("Embarked"),
  myDummyCode("Title")[,-5],
  myDummyCode("Surname")[,-1],
  myDummyCode("TicketPrefix")[,-1],
  # worse: myDummyCode("CabinSideOrZero")[,c("CabinSideOrZero_0", "CabinSideOrZero_2")],
  myDummyCode("DeckOrUnknown")
)

cv.lasso = cv.glmnet(as.matrix(lasso.X[ITrain,]), as.matrix(lasso.y[ITrain]), alpha=1)
plot(cv.lasso)
best.lambda <- cv.lasso$lambda.min
print(best.lambda)

lasso.model <- glmnet(as.matrix(lasso.X[ITrain,]), as.matrix(lasso.y[ITrain]), alpha=1, lambda=best.lambda*3)
lasso.coefs <- coef(lasso.model)
lasso.selected <- rownames(lasso.coefs)[as.numeric(lasso.coefs) != 0.0]

lasso.coefs

lasso.f <- as.formula(
  paste("Survived ~ ", paste(ifelse(lasso.selected == "(Intercept)", "1", lasso.selected), collapse = " + "))
)

lasso.gam_df <- cbind(Survived=lasso.y, all_df[, c("kfold", "PassengerId")], lasso.X)

evaluate_gam_formula(lasso.f, lasso.gam_df[ITrain,])

write_gam_solution(lasso.f, lasso.gam_df, "lasso.solution.csv") # 0.78468

```

## PCA

```{r}

pca.data <- scale(lasso.X)
pca.components <- princomp(pca.data)
summary(pca.components)
plot(pca.components$sdev)

for(i in seq(5)) {
  tmp_loading <- pca.components$loadings[, i]
  tmp_loading <- tmp_loading[order(abs(tmp_loading), decreasing = TRUE)[1:5]]
  print(data.frame(setNames(list(tmp_loading), paste0("pca", i))))
  
  print(
    ggplot(data.frame(pca = pca.components$scores[ITrain,i], Survived = factor(lasso.y[ITrain])), aes(x=pca)) +
      geom_density(aes(group=Survived, col=Survived))
  )
}

```

## ICA

```{r}

# install.packages("fastICA")
library(fastICA)

ica.data <- lasso.X[ITrain,]
ica.estimation <- fastICA(ica.data, 5, alg.typ = "parallel", fun = "logcosh", alpha = 1, method = "C", row.norm = FALSE, maxit = 200, tol = 0.0001, verbose = TRUE)

for(i in seq(5)) {
  print(
    ggplot(data.frame(ica = ica.estimation$S[,i], Survived = lasso.y[ITrain]), aes(x=ica)) +
      geom_density(aes(group=Survived, col=Survived))
  )
}

```

## r-TSNE

```{r}
# install.packages("Rtsne")
library(Rtsne)

# Warum haben wir Dubletten??
# duplicates <- rbind((lasso.X[ITrain,])[duplicated(lasso.X[ITrain,]),], (lasso.X[ITrain,])[duplicated(lasso.X[ITrain,], fromLast=TRUE),])
# duplicates <- duplicates[order(duplicates$AgeImputed, duplicates$SibSp, duplicates$Parch),]
# duplicates
# all_df[rownames(duplicates),]

tsne.data <- unique(lasso.X[ITrain,])
tsne.results <- Rtsne(tsne.data, dims = 2, perplexity = 25, verbose = TRUE, max_iter = 1500)
tsne.df <- data.frame(
  X = tsne.results$Y[, 1],
  Y = tsne.results$Y[, 2],
  Sex = tsne.data$Sex,
  Pclass = tsne.data$Pclass,
  Survived = all_df[rownames(tsne.data),"Survived"]
)
colors <- c("#E6194B", "#3CB44B")

ggplot(tsne.df, aes(x = X, y = Y, color = factor(Survived, levels=c(1,2)))) +
  geom_point(size = 1.5) +
  scale_color_manual(values = colors) +
  labs(
    title = "t-SNE 2-Dimensional Digit Visualization",
    x = "t-SNE Dimension 1",
    y = "t-SNE Dimension 2"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20)
  )

```

## Random Forest

```{r}

library(caret)

ctrl <- rfeControl(
  functions = lmFuncs, #rfFuncs,
  method = "cv",
  number = 5
)

rfe_fit <- rfe(
  x = lasso.X[ITrain,],
  y = lasso.y[ITrain],
  sizes = c(5, 15, 30, 35, 40, 45, 50, 55, 75),
  rfeControl = ctrl
)

rfe_fit$results

rfe_fit$evaluation <- data.frame()
for(s in rfe_fit$results$Variables ) {
  rfe.attributes <- unique(rfe_fit$variables[rfe_fit$variables$Variables == s,]$var)
  rfe.attributes
  rfe.f <- as.formula(
    paste("Survived ~ ", paste(rfe.attributes, collapse = " + "))
  )
  rfe.gam_df <- cbind(Survived=lasso.y, all_df[, c("kfold", "PassengerId")], lasso.X)
  
  rfe_fit$evaluation <- rbind(rfe_fit$evaluation, evaluate_gam_formula(rfe.f, rfe.gam_df[ITrain,]))
  
  write_gam_solution(rfe.f, rfe.gam_df, paste0("rfe.",s,".solution.csv"))
}
rfe_fit$evaluation

# rfe.35.solution.csv: 0.77751
```


## Boruta

```{r}

library(Boruta)

set.seed(123)

boruta <- Boruta(
  reformulate(colnames(lasso.X), response = "Survived"),
  data = cbind(Survived = lasso.y, lasso.X)[ITrain,],
  doTrace = 1
)

boruta
boruta.attributes <- getSelectedAttributes(boruta, withTentative = FALSE)
boruta.f <- as.formula(
  paste("Survived ~ ", paste(boruta.attributes, collapse = " + "))
)

boruta.gam_df <- cbind(Survived=lasso.y, all_df[, c("kfold", "PassengerId")], lasso.X)
evaluate_gam_formula(boruta.f, boruta.gam_df[ITrain,])
write_gam_solution(boruta.f, boruta.gam_df, "boruta.solution.csv")


```

# Baseline Models

Here we use all the simple Features in a bayesian regression model to establish a baseline:

## Baseline Rnadom Forest

```{r}

# install.packages("ranger")
library(ranger)

set.seed(123)

rf_fit <- ranger(
  Survived ~ Pclass + AgeImputed + Fare + LogFare + Sex + SibSp + Parch + SoloTraveler + AdultWithInfant + AdultWithBoy + AdultWithTeenager + AdultWithOld + TicketTravelers + Namesakes +
    AgeBin + FareBin + Embarked + Title + Surname + TicketPrefix,
  data = all_df[ITrain,],
  num.trees = 500,
  mtry = floor(sqrt(ncol(all_df) - 1)),
  min.node.size = 5,
  probability = TRUE,          # for class probabilities
  importance = "permutation"
)
imp <- sort(rf_fit$variable.importance, decreasing = TRUE)
barplot(
  imp,
  horiz = TRUE,
  las = 1,
  main = "Permutation importance"
)
data.frame(imp)

```

## Baseline 1
```{r}
baseline1_df <- all_df
baseline1_df$Title <- as.character(baseline1_df$Title)
baseline1_df[is.na(baseline1_df$Title) | baseline1_df$Title == "Dr" | baseline1_df$Title == "Col" | baseline1_df$Title == "Major" | baseline1_df$Title == "Mlle" | baseline1_df$Title == "Rev" | baseline1_df$Title == "Miss" | baseline1_df$Title == "Mrs",]$Title <- "NoTitle"

baseline1_formula <- bf(Survived ~ Pclass + Embarked + AgeImputed + Sex + Title + SibSp + Parch + Fare) # TicketPrefix & Deck & Surname have NAs
baseline1_fit <- brm(
  baseline1_formula,
  data = baseline1_df[ITrain,],
  family = bernoulli(),
  iter = 2000,
  silent = FALSE,
  refresh = 100
)
baseline1_evaluation <- evaluate_brms_model(baseline1_fit)
baseline1_fit
baseline1_evaluation

baseline1_prediction <- predict(baseline1_fit, newdata=baseline1_df[ITest,])
baseline1_solution <- data.frame(PassengerId = all_df[ITest,]$PassengerId, Survived=(baseline1_prediction[,"Estimate"] >= 0.5) * 1)

write.csv(baseline1_solution, file = "baseline1.solution.csv", row.names = FALSE) # Test-Set: 0.76555
```

## Baseline 2

```{r}
baseline2_df <- baseline1_df
baseline2_df$Deck <- as.character(baseline2_df$Deck)
baseline2_df[is.na(baseline2_df$Deck) | baseline2_df$Deck == "G" | baseline2_df$Deck == "T",]$Deck <- "NoDeck"

baseline2_formula <- bf(Survived ~ Pclass + Embarked + AgeImputed + Sex + Title + SibSp + Parch + Fare + Deck)
baseline2_fit <- brm(
  baseline2_formula,
  data = baseline2_df[ITrain,],
  family = bernoulli(),
  iter = 2000
)
baseline2_evaluation <- evaluate_brms_model(baseline2_fit)
baseline2_fit
baseline2_evaluation

baseline2_prediction <- predict(baseline2_fit, newdata=baseline2_df[ITest,])
baseline2_solution <- data.frame(PassengerId = all_df[ITest,]$PassengerId, Survived=(baseline2_prediction[,"Estimate"] >= 0.5) * 1)

write.csv(baseline2_solution, file = "baseline2.solution.csv", row.names = FALSE) # Test-Set: 0.75598
```

## Baseline 3

```{r}
baseline3_df <- baseline2_df

#baseline3_formula <- bf(Survived ~ Pclass + Embarked + AgeBin*Sex + Title + SibSp + Parch + FareBin + Deck)
baseline3_formula <- bf(Survived ~ Pclass + Embarked + AgeBin*Sex + Title + SibSp + Parch + LogFare + Deck)
baseline3_fit <- brm(
  baseline3_formula,
  data = baseline3_df[ITrain,],
  family = bernoulli(),
  iter = 2000
  # max_treedepth = 15
)
baseline3_evaluation <- evaluate_brms_model(baseline3_fit)
baseline3_fit
baseline3_evaluation

baseline3_prediction <- predict(baseline3_fit, newdata=baseline3_df[ITest,])
baseline3_solution <- data.frame(PassengerId = all_df[ITest,]$PassengerId, Survived=(baseline3_prediction[,"Estimate"] >= 0.5) * 1)

write.csv(baseline3_solution, file = "baseline3.solution.csv", row.names = FALSE) # Test-Set: 0.74401
```

## Baseline 4 (GPs, TPs)

```{r}
baseline4_df <- baseline3_df
baseline4_df[,c("Pclass", "Sex")] <- lapply(baseline4_df[,c("Pclass", "Sex")], FUN=as.numeric)
baseline4_df[,c("Deck")] <- as.numeric(all_df$Deck)
baseline4_df[is.na(baseline4_df$Deck),"Deck"] <- mean(baseline4_df$Deck, na.rm = TRUE)
baseline4_df$Title <- as.numeric(factor(baseline4_df$Title))

results <- rbind(
  #evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, bs = "tp"), baseline4_df),
  #evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "tp"), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, LogFare, bs = "tp"), baseline4_df),
  #evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, bs = "tp", k=400), baseline4_df),
  #evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "tp", k=400), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, LogFare, bs = "tp", k=400), baseline4_df)
)
results

results <- rbind(
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, bs = "gp", k = 50), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, bs = "gp", k = 50), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 50), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, LogFare, bs = "gp", k = 50), baseline4_df)
)
results

results <- rbind(
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 40), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 50), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 60), baseline4_df), # Best
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 70), baseline4_df),
  evaluate_gam_formula(Survived ~ s(Pclass, AgeImputed, Sex, SibSp + Parch, Deck, bs = "gp", k = 80), baseline4_df)
)
results

#baseline3_formula <- bf(Survived ~ Pclass + Embarked + AgeBin*Sex + Title + SibSp + Parch + FareBin + Deck)
# baseline4_formula <- bf(Survived ~ gp(Pclass, AgeImputed, Sex, k = 15)) # , LogFare, Deck, iso = FALSE
# baseline4_fit <- brm(
#   baseline4_formula,
#   data = baseline4_df[ITrain,],
#   family = bernoulli(),
#   iter = 2000
# )
# baseline4_evaluation <- evaluate_brms_model(baseline3_fit)
# baseline4_fit
# baseline4_evaluation
```
# The End
